### 2017-03-10 real/draft.tex
[draft] Editing unbinned section

Let's consider the extreme case $\npack \to \infty$ so we can neglect
the uncertainty on $\alpha, \beta$ and
$p(\alpha, \beta \cond \npack, \bml) =
\delta(\alpha-\alpha_0)\delta(\beta - \beta_0)$ such that
\begin{align}
  \lleq{gmj}
  p(Q\cond n, \alpha_0, \beta_0)
  &=     \sum_{N=0}^\infty p(N\cond n) p(Q\cond N, \alpha_0, \beta_0) \, .
\end{align}
For illustration only, let us consider the Poisson rate $\lambda$ were
known. Then learning $n$ would not improve our knowledge of $\lambda$
and we would have
\begin{align}
  \lleq{gmk}
  p(Q\cond \lambda, \alpha_0, \beta_0)
  &=     \sum_{N=0}^\infty p(N\cond \lambda) p(Q\cond N, \alpha_0, \beta_0) \, ,
\end{align}
where $p(N \cond \lambda)$ is the Poisson distribution \refeq{pmb}.
In \reffig{comp-unc}, we illustrate the impact of the uncertainty due
to $\lambda$ or $\alpha, \beta$ for $n=0,2, 20$ packets and continue
to use as true values $\alpha_0 = 1.5, \beta_0 = 60$. For $n>0$, we
set $\lambda=n$ and for $n=0$, we set $\lambda = 0.347$ by solving
\begin{align}
  \lleq{gml}
  p(N=0 \cond \lambda) &= p(N=0 \cond n=0)
\end{align}
such that the contribution at $N=0$ is identical for the Poisson
likelihood and Poisson posterior predictive distribution.

@@

Already for $n=20$, it becomes apparent in the right panel of
\reffig{comp-unc} that the uncertainty for $\Lumtot$ is dominated by
the Poisson uncertainty: both $p(Q \cond n, \alpha_0, \beta_0)$ and
$p(Q \cond n, \bml)$ yield a standard error of $\approx 0.19$.

The case where all parameters are known is not attainable in practice
and only shown for comparison. But the case where $\alpha, \beta$ are
fixed can be reached easily if many packets from other bins can be
combined. Note that $n=2$ is the minimum number of packets to perform
the binned analysis.

@@

While the integral in \refeq{ral} cannot be integrated analytically,
the integrand is unimodal and strongly peaked, so that the Laplace or
saddle-point approximation~\cite[Ch. 27]{mackay2003information} can be
applied. If only the value $p(\Lumtot{=}\lumtot \cond n,m_1,m_2)$ at the
total data luminosity $q$ is sought, then one can approximate it
directly. In addition, we can compute the moments of $\Lumtot$ through
the \emph{moment generating function} \cite{stuart1994kendall}
\begin{align}
  \lleq{rap}
  M(t \cond \bmy) = \int_0^{\infty} \rmdx{\Lumtot} p(\Lumtot \cond \bmy) e^{-t \Lumtot} \,
\end{align}
from which the moments are accessible by differentiation
\begin{align}
  \lleq{raq}
  \expect{\Lumtot^k} = (-1)^k \eval{\dv[k]{M(t \cond \bmy)}{t}}_{t=0}
\end{align}
Assuming all involved integrals are absolutely convergent, we can
exchange the order of integration and find
\begin{align}
  \lleq{rar}
  M(t \cond \bmy) &= \int \rmdx{\lambda} \rmdx{\mu} \rmdx{\sigma^2} p(\lambda \cond \bmy) p( \mu, \sigma^2 \cond \bmy)  f(t \cond \lambda, \mu, \sigma^2),
  \intertext{where }
  f\left(t \cond  \lambda, \mu, \sigma^2\right) &\equiv \int \rmdx{\Lumtot} \GaussianDist \left( \Lumtot \cond \lambda \mu, \lambda (\mu^2 + \sigma^2) \right) e^{-t \Lumtot}\\
  &= \frac{1}{2} \exp \qty(t/2 \lambda (t (\mu^2 + \sigma^2)- 2\mu)) \erfc \qty( \frac{\sqrt{\lambda}(t \qty(\mu^2 + \sigma^2) - \mu)}{\sqrt{2 \qty(\mu^2 + \sigma^2)}}).
\end{align}
This implies that the mean and variance of $\Lumtot$ require an integral over the three parameters
\begin{align}
  \lleq{ras}
  \expect{\Lumtot} &= \int \rmdx{\lambda} \rmdx{\mu} \rmdx{\sigma^2} p(\lambda \cond n) p( \mu, \sigma^2 \cond \bmy) \qty(-\eval{\dv{f(t \cond  \lambda, \mu, \sigma^2)}{t}}_{t=0})\\
  \variance{\Lumtot} &= \int \rmdx{\lambda} \rmdx{\mu} \rmdx{\sigma^2} p(\lambda \cond n) p( \mu, \sigma^2 \cond \bmy) \qty{\eval{\dv[2]{f(t \cond  \lambda, \mu, \sigma^2)}{t}}_{t=0} - \qty(\eval{\dv{f(t \cond  \lambda, \mu, \sigma^2)}{t}}_{t=0})^2} \, .
\end{align}
\fred{We need a plot, the one to compare predictions and 250 real replicas or 100 virtual replicas?} %
\fred{Advantage of this approach: no need to store all packets anymore, sum and sum of squares can be accumulated online $\Rightarrow$ reduced computation and storage requirements}

\section{Delendum}

\subsection{Numerical implementation} \label{sec:numerical}

\fred{Editing will be done by Fred}

In our numerical implementation, the constants
$(\alpha_{\rm max}, \beta_{\rm max})$ are chosen by trial and error to
be large enough to contain the relevant region of
$p(\ell\cond \alpha,\beta)$.
%
To perform the integrals over $(\alpha,\beta)$ to find $C$ and the
Gamma-weighted numerator in (\ref{spe}), we used the fact that the
posterior $p(\alpha,\beta \cond n,\bml)$ is approximately Gaussian in
both $\alpha$ and $\beta$ for large $n$, so we can apply Laplace's
method for both. This requires the Hessian at the mode and the
gradient helps in locating the mode. The relevant analytic expressions
are summarized in \refsec{appendix}.

Regarding the sum over $N$, we start at
$N = \lfloor n-a+1 \rfloor = \arg \max_N p(N|n)$ %
\fred{Actually we should say $p(N \cond n, a)$} %
and continue to larger and smaller values of $N$ until the
contribution to $p(Q \cond n,\bml)$ becomes negligible. In case the
relevant contribution is from large $N$ and $n$, we interpret $N$ as a
continuous variable, replace factorials by $\Gamma$ function in
\refeq{spj}, and apply Laplace's method to integrate over $N$ also.

>>>>

Let's consider the extreme case $\npack \to \infty$ so we can neglect
the uncertainty on $\alpha, \beta$ and
$p(\alpha, \beta \cond \npack, \bml) =
\delta(\alpha-\alpha_0)\delta(\beta - \beta_0)$ such that
\begin{align}
  \lleq{gmj}
  p(Q\cond n, \alpha_0, \beta_0)
  &=     \sum_{N=0}^\infty p(N\cond n) p(Q\cond N, \alpha_0, \beta_0) \, .
\end{align}
For illustration only, let us consider the Poisson rate $\lambda$ were
known. Then learning $n$ would not improve our knowledge of $\lambda$
and we would have
\begin{align}
  \lleq{gmk}
  p(Q\cond \lambda, \alpha_0, \beta_0)
  &=     \sum_{N=0}^\infty p(N\cond \lambda) p(Q\cond N, \alpha_0, \beta_0) \, ,
\end{align}
where $p(N \cond \lambda)$ is the Poisson distribution \refeq{pmb}.
In \reffig{comp-unc}, we illustrate the impact of the uncertainty due
to $\lambda$ or $\alpha, \beta$ for $n=0,2, 20$ packets and continue
to use as true values $\alpha_0 = 1.5, \beta_0 = 60$. For $n>0$, we
set $\lambda=n$ as the mode of the posterior $p(\lambda \cond n)$, and
for $n=0$, we set $\lambda = 0.347$ by solving
\begin{align}
  \lleq{gml}
  p(N=0 \cond \lambda) &= p(N=0 \cond n=0)
\end{align}
such that the contribution at $N=0$ is identical for the Poisson
likelihood and Poisson posterior predictive distribution.

@@

Already for $n=20$, it becomes apparent in the right panel of
\reffig{comp-unc} that the uncertainty for $\Lumtot$ is dominated by
the Poisson uncertainty: both $p(Q \cond n, \alpha_0, \beta_0)$ and
$p(Q \cond n, \bml)$ yield a standard error of $\approx 0.19$. This is
necessarily larger that the standard error of
$p(\Lumtot \cond \lambda, \alpha_0, \beta_0)$ because the variance of

The case where all parameters are known is not attainable in practice
and only shown for reference. But the case where $\alpha, \beta$ are
fixed can be approximated easily if many packets from other bins can
be combined.

Note that $n=2$ is the minimum number of packets to perform
the binned analysis.
\fred{I was playing around in mathematica and it seems even for $n=1$ the posterior has a well defined mode. I don't if it is proper, though. In any case it has huge variance.}

@@

While the integral in \refeq{ral} cannot be integrated analytically,
the integrand is unimodal and strongly peaked, so that the Laplace or
saddle-point approximation~\cite[Ch. 27]{mackay2003information} can be
applied. If only the value $p(\Lumtot{=}\lumtot \cond n,m_1,m_2)$ at the
total data luminosity $q$ is sought, then one can approximate it
directly. In addition, we can compute the moments of $\Lumtot$ through
the \emph{moment generating function} \cite{stuart1994kendall}
\begin{align}
  \lleq{rap}
  M(t \cond \bmy) = \int_0^{\infty} \rmdx{\Lumtot} p(\Lumtot \cond \bmy) e^{-t \Lumtot} \,
\end{align}
from which the moments are accessible by differentiation
\begin{align}
  \lleq{raq}
  \expect{\Lumtot^k} = (-1)^k \eval{\dv[k]{M(t \cond \bmy)}{t}}_{t=0}
\end{align}
Assuming all involved integrals are absolutely convergent, we can
exchange the order of integration and find
\begin{align}
  \lleq{rar}
  M(t \cond \bmy) &= \int \rmdx{\lambda} \rmdx{\mu} \rmdx{\sigma^2} p(\lambda \cond \bmy) p( \mu, \sigma^2 \cond \bmy)  f(t \cond \lambda, \mu, \sigma^2),
  \intertext{where }
  f\left(t \cond  \lambda, \mu, \sigma^2\right) &\equiv \int \rmdx{\Lumtot} \GaussianDist \left( \Lumtot \cond \lambda \mu, \lambda (\mu^2 + \sigma^2) \right) e^{-t \Lumtot}\\
  &= \frac{1}{2} \exp \qty(t/2 \lambda (t (\mu^2 + \sigma^2)- 2\mu)) \erfc \qty( \frac{\sqrt{\lambda}(t \qty(\mu^2 + \sigma^2) - \mu)}{\sqrt{2 \qty(\mu^2 + \sigma^2)}}).
\end{align}
This implies that the mean and variance of $\Lumtot$ require an integral over the three parameters
\begin{align}
  \lleq{ras}
  \expect{\Lumtot} &= \int \rmdx{\lambda} \rmdx{\mu} \rmdx{\sigma^2} p(\lambda \cond n) p( \mu, \sigma^2 \cond \bmy) \qty(-\eval{\dv{f(t \cond  \lambda, \mu, \sigma^2)}{t}}_{t=0})\\
  \variance{\Lumtot} &= \int \rmdx{\lambda} \rmdx{\mu} \rmdx{\sigma^2} p(\lambda \cond n) p( \mu, \sigma^2 \cond \bmy) \qty{\eval{\dv[2]{f(t \cond  \lambda, \mu, \sigma^2)}{t}}_{t=0} - \qty(\eval{\dv{f(t \cond  \lambda, \mu, \sigma^2)}{t}}_{t=0})^2} \, .
\end{align}
\fred{Advantage of this approach: no need to store all packets anymore, sum and sum of squares can be accumulated online $\Rightarrow$ reduced computation and storage requirements}
